SOMETHING ABOUT THE DISTRIBUTED AVERAGING BASED APPROXIMATION IN SENSOR NETWORKS

### 2022-10-14 paper.tex
smaller font size for more space

\begin{document}
% \ninept
%
\maketitle
%
\begin{abstract}
  Distributed signal-processing algorithms in (wireless) sensor networks often aim to decentralize processing tasks to reduce transmission cost and computational complexity or avoid reliance on a single device for processing. In this contribution, we extend a distributed adaptive algorithm for blind system identification that relies on the knowledge of a stacked network-wide solution vector at each node, the computation of which requires either broadcasting or relaying of node-specific values to all other nodes. The extended algorithm employs a distributed-averaging-based estimation scheme to estimate the network-wide norm value by only using the information provided by neighboring sensor nodes. We introduce a mixing factor between instantaneous and recursion values for adaptivity in a time-varying system. The extension leads to a decreased number of transmissions while maintaining estimation performance close to the optimal case. Simulation results show how the mixing factor and the averaging scheme's number of iterations influence performance.
\end{abstract}
%
\begin{keywords}
multi-channel signal processing, distributed signal processing, wireless sensor networks, blind system identification, distributed averaging
\end{keywords}
%
\section{INTRODUCTION}
\label{sec:intro}

@@

\subsection[]{Sensor network}
We assume a set \(\Mset \triangleq \{1,\ldots,M\}\) sensor nodes and a set of edges \(\mathcal{E}\) which connect the nodes forming a sensor network.
Each edge is an unordered pair of nodes \(\{i,j\} \in \mathcal{E}\), which represents the communication link between them.
From this follows the neighborhood of a node \(i\), i.e., the other nodes it is directly connected to, defined as \(\Nset_i = \{j|\{i,j\} \in \mathcal{E}\}\).
It is to note, that we define the set with \(i \in \Nset_i\) for ease of notation later.
This does however not represent an actual inter-node link.
Furthermore, we can define the symmetric adjacency matrix \(\mtxb{C}\) with elements
\begin{equation}
    C_{ij} = \begin{cases}
        1 \quad \text{if } i = j\\
        1 \quad \text{if } i \neq j \text{ and }\{i,j\} \in \mathcal{E}\\
        0 \quad \text{otherwise}.
    \end{cases}
\end{equation}
It is to note that in \cite{blochbergerDBSI}, the pairs \(\{i,j\} \in \mathcal{E}\) are ordered, which yields a directed graph and leads to two sets of neighborhoods ("transmit" and "receive") and therefore a non-symmetric adjacency matrix.
In this paper, we limit the explanation to the undirected graph to save space, as the extension is straightforward.

@@

\begin{algorithm}[t]
    \caption{ADMM BSI with FDLA-based adaptive estimation of norm values}\label{alg:davg_norm_est}
    \(\W \gets\) \eqref{eq:adaptivenormest:fdlaminprob}\;
    \(\bar{\phi}_i^{0} \gets 1, \forall i \in \Mset\)\;
    \For(){\(m=1\dots\)}
    {
        \For(){\(i \in \Mset\)}
        {
            % \(\cdots\)\\
            % \(\w_i^m \gets \operatorname{argmin}_{\w_i}\, \mathcal{L}_{\rho} (\w,\h^m,\uu^m)\)\;
            % Compute \(\bar{\h}_i^{m}\)\;
            % \(\cdots\)\\
            \emph{The steps before are as introduced in }\cite{blochbergerDBSI}\\
            \dotfill\\
            Transmit \(\bar{\h}_i^{m}\) to nodes  \(j \in \Nset_i\)\;
            \(\bar{\phi}_i^{m} \gets \sum_{j \in \Nset_i} W_{ij} \|\bar{\h}_i^m\|^2\)\;
            \(\gamma_i^{m} \gets \min \left\lbrace \frac{| \bar{\phi}_i^{m} - \bar{\phi}_i^{m-1} |}{\bar{\phi}_i^{m-1}},\,1\right\rbrace\)\;
            \eIf(){\(m = 1\)}
            {
                \(\phi_i^{m}(1) \gets \|\bar{\h}_i^m\|^2\)\;
            }
            {
                \(\phi_i^{m}(1) \gets \phi_i^{m-1}(K)\)\;
            }
            \For(){\(k=1,\dots,K\)}
            {
                Transmit \(\phi_i(k)\) to nodes  \(j \in \Nset_i\)\;
                \(\phi_i({k+1}) \gets \sum_{j \in \Nset_i} W_{ij} \phi_j({k})\)\;
            }
            \(\hat{\phi}_i^{m} \gets \gamma_i \bar{\phi}_i^{m} + (1-\gamma) \phi_i^{m}(K)\)\;
            \(\hat{\h_i}^{m} \gets \frac{\bar{\h}_i^{m}}{\sqrt{\hat{\phi}_i^{m}M}}\)\;
            \dotfill\\
            \emph{The steps after are as introduced in }\cite{blochbergerDBSI}\\
            % \(\cdots\)\\
            % \(\uu_i^{m} \gets \uu_i^{m-1} + \rho \left( \w_i^{m} - \tilde{\h}_i^{m} \right)\)\;
            % \(\cdots\)\\
            % % \,\\
        }
    }
\end{algorithm}

@@

\section{Evaluation}
\label{sec:simulations}
\subsection[]{Transmission cost}
\label{sec:transcost}
\begin{table}[t]
    \centering
    \begin{tabular}{ |l|l|l| } 
        \hline
        & Transmit & Receive \\
        \hline\hline
        (1) Full & \(\mathcal{O}(M-1)\) & \(\mathcal{O}(M-1)\) \\
        \hline
        (2) Broadcast & \(\mathcal{O}(1)\) & \(\mathcal{O}(M-1)\) \\ 
        \hline
        (3) Neigborhood & \(\mathcal{O}(\bar{N}_i K)\) & \(\mathcal{O}(\bar{N}_i K)\) \\ 
        \hline
    \end{tabular}
    \caption[]{Comparison of communication cost.}
    \label{tab:transcost:table}
\end{table}
To describe the communication cost within the network, we adopt the Big-\(\mathcal{O}\) notation, similar to the complexity analysis of algorithms.
We consider the communication of a variable from one node to another as \(\mathcal{O}(1)+\mathcal{O}(1)\), where the two terms represent transmit and receive operations, respectively, as both require computational resources at a node.
The most useful measure in the case of this work is to analyze the cost per node.
For this we define an average neighborhood size \(\bar{N}_i = \frac{1}{M} \sum_{i \in \Mset} N_i\) where \(N_i = | \Nset_i |\) is the cardinality/size of the neigborhood set for node \(i\).
We compare the following communication schemes:
\begin{itemize}
    \itemsep-0.2em
    \item[(1)] A fully connected network, i.e., all nodes are linked with all other nodes by direct links, \(\Nset_i = \Mset\).
    \item[(2)] The nodes employ broadcasting, i.e., the nodes transmit to all other nodes without direct links.
    \item[(3)] The nodes only communicate with neighboring nodes, \(\Nset_i \subset \Mset\).
\end{itemize}
For (1), the cost increases quadratically with increasing network size \(M\), both in transmit and receive operations, whereas for (2), the quadratic increase happens only at the latter.
Using communication scheme (3), as in \cite{yuDistributedBlindSystem2014,liuDistributedBlindIdentification2016} and proposed in this paper, cost in proportional to neighborhood size and iteration count \(K\).
Refer to \autoref{tab:transcost:table} for the collected orders of communication cost.

@@

\subsection[]{Simulation}
\begin{figure}[t]
    \centering
    \input{simulations/plots/icassp2023-static-time/icassp2023-static-time.pgf}
    \caption[]{Median NPM of 30 randomized runs over time for optimal case where all node information is available (fully connected network/broadcasting), fixed values of \(\gamma_i\) with \(K \in \{1,10\}\), adaptive \(\gamma_i\) with \(K=1\). The adaptive method yields results very close to the optimal method. A moving average filter was applied to result curves for better readability (compare transparent and opaque lines).}
    \label{fig:simulations:NPMtime}
\end{figure}
\begin{figure}[t]
    \centering
    \input{simulations/plots/icassp2023-static/icassp2023-static.pgf}
    \caption[]{Mean over time of median NPM of 30 randomized runs after convergence for varied values for fixed \(\gamma_i\) on the interval \([0.0, 0.04]\) and \(K \in \{1,2,10\}\). Results for optimal method, and adaptive \(\gamma_i\) included for comparison.}
    \label{fig:simulations:avgNPMgamma}
\end{figure}
To evaluate the effectiveness of the proposed method, simulations were run.
For this we define the error measure, the normalized projection misalignment
\begin{equation}
    \begin{aligned}
        \text{NPM}(m) &= 20\,\log_{10} \frac{\left\| \mtxb{e} \right\|}{\left\|\h(m)\right\|},\\
        \text{with}\quad \mtxb{e} &= \h(m) - \frac{\h^\T (m) \h_{\text{t}}}{\h_{\text{t}}^\T \h_{\text{t}}}\h(m),
    \end{aligned}
\end{equation}
commonly used in BSI to compare estimated \(\h\) and true \(\h_t\).
The simulation setup is a network with \(M=5\) nodes arranged in a ring topology, each node having \(|N_i|=2\) neighbors.
The input signal is zero-mean white Gaussian noise (WGN), the impulse responses of length \(L=16\) are zero-mean WGN as well, and at each channel, independent WGN is added with \(\text{SNR}=10\,\text{dB}\).
The norms of the impulse responses are scaled to random values from the uniform distribution \(\mathcal{U}_{[0.5,2.0]}\).
A comparison is made between the following cases:
\begin{itemize}
    \itemsep-0.2em
    \item[(a)] all information is available for global norm computation ("optimal" in \autoref{fig:simulations:NPMtime} \& \autoref{fig:simulations:avgNPMgamma}),
    \item[(b)] neighborhood communication with fixed \(\gamma_i \in [0.0, 0.4]\) and \(K \in \{1,2,10\}\),
    \item[(b)] neighborhood communication with adaptive \(\gamma_i\) \eqref{eq:adaptivenormest:adaptivegamma} and \(K=1\).
\end{itemize}
\autoref{fig:simulations:NPMtime} shows the median of 30 Monte-Carlo runs.
We can observe that when fixing \(\gamma_i\), a tradeoff between convergence speed and steady-state error arises at low iteration counts (cf. also \autoref{fig:simulations:avgNPMgamma}).
With \(K=10\), the performance comes close to the optimal case at the cost of additional communication between nodes.
Compare that to the case with adaptive \(\gamma_i\) where even at \(K=1\), i.e., no additional communication for the recursive estimation scheme, convergence speed and steady-state error are close to the optimal case.

\begin{todo}
    Now that I write this, i wonder if evaluating this extension would rather revolve around measuring how well the norms are approximated\dots
\end{todo}
\begin{todo}
    evaluate on time-varying norms of IRs if there is space
\end{todo}

>>>>

\begin{document}
\ninept
%
\maketitle
%
\begin{abstract}
    Distributed signal-processing algorithms in (wireless) sensor networks often aim to decentralize processing tasks to reduce communication cost and computational complexity or avoid reliance on a single device for processing.
    In this contribution, we extend a distributed adaptive algorithm for blind system identification that relies on the knowledge of a stacked network-wide solution vector at each node, the computation of which requires either broadcasting or relaying of node-specific values to all other nodes.
    The extended algorithm employs a distributed-averaging-based estimation scheme to estimate the network-wide norm value by only using the information provided by neighboring sensor nodes.
    We introduce an adaptive mixing factor between instantaneous and recursion values for adaptivity in a time-varying system.
    Simulation results show that the extension provides estimation results close to the optimal fully-connected-network case while reducing inter-node transmission significantly.
\end{abstract}
%
\begin{keywords}
multi-channel signal processing, distributed signal processing, wireless sensor networks, blind system identification, distributed averaging
\end{keywords}
%
\section{INTRODUCTION}
\label{sec:intro}

@@

\subsection[]{Sensor network}
We assume a set \(\Mset \triangleq \{1,\ldots,M\}\) sensor nodes and a set of edges \(\mathcal{E}\) which connect the nodes forming a sensor network.
Each edge is an unordered pair of nodes \(\{i,j\} \in \mathcal{E}\), which represents the communication link between them.
From this follows the neighborhood of a node \(i\), i.e., the other nodes it is directly connected to, defined as \(\Nset_i = \{j|\{i,j\} \in \mathcal{E}\}\).
It is to note, that we define the set with \(i \in \Nset_i\) for ease of notation later.
This does however not represent an actual inter-node link.
Furthermore, we can define the symmetric adjacency matrix \(\mtxb{C}\) with elements
\begin{equation}
    C_{ij} = \begin{cases}
        1 \quad \text{if } i = j\\
        1 \quad \text{if } i \neq j \text{ and }\{i,j\} \in \mathcal{E}\\
        0 \quad \text{otherwise}.
    \end{cases}
\end{equation}
It is to note that in \cite{blochbergerDBSI}, the pairs \(\{i,j\} \in \mathcal{E}\) are ordered, which yields a directed graph and leads to two sets of neighborhoods ("transmit" and "receive") and therefore a non-symmetric adjacency matrix.
For the sake of brevity, we limit the explanations in this paper to the undirected case.

@@

\begin{algorithm}[t]
    \caption{ADMM BSI with FDLA-based adaptive estimation of norm values}\label{alg:davg_norm_est}
    \(\W \gets\) \eqref{eq:adaptivenormest:fdlaminprob}\;
    \(\bar{\phi}_i^{0} \gets 1, \forall i \in \Mset\)\;
    \For(){\(m=1\dots\)}
    {
        \For(){\(i \in \Mset\)}
        {
            % \(\cdots\)\\
            % \(\w_i^m \gets \operatorname{argmin}_{\w_i}\, \mathcal{L}_{\rho} (\w,\h^m,\uu^m)\)\;
            % Compute \(\bar{\h}_i^{m}\)\;
            % \(\cdots\)\\
            \emph{The steps before are as introduced in }\cite{blochbergerDBSI}\\
            \dotfill\\
            Transmit \(\bar{\h}_i^{m}\) to nodes  \(j \in \Nset_i\)\;
            \(\bar{\phi}_i^{m} \gets \sum_{j \in \Nset_i} W_{ij} \|\bar{\h}_i^m\|^2\)\;
            \(\gamma_i^{m} \gets \min \left\lbrace \frac{| \bar{\phi}_i^{m} - \bar{\phi}_i^{m-1} |}{\bar{\phi}_i^{m-1}},\,1\right\rbrace\)\;
            \eIf(){\(m = 1\)}
            {
                \(\phi_i^{m}(1) \gets \|\bar{\h}_i^m\|^2\)\;
            }
            {
                \(\phi_i^{m}(1) \gets \hat{\phi}_i^{m-1}(K)\)\;
            }
            \For(){\(k=1,\dots,K\)}
            {
                Transmit \(\phi_i(k)\) to nodes  \(j \in \Nset_i\)\;
                \(\phi_i({k+1}) \gets \sum_{j \in \Nset_i} W_{ij} \phi_j({k})\)\;
            }
            \(\hat{\phi}_i^{m} \gets \gamma_i \bar{\phi}_i^{m} + (1-\gamma) \phi_i^{m}(K)\)\;
            \(\hat{\h_i}^{m} \gets \frac{\bar{\h}_i^{m}}{\sqrt{\hat{\phi}_i^{m}M}}\)\;
            \dotfill\\
            \emph{The steps after are as introduced in }\cite{blochbergerDBSI}\\
            % \(\cdots\)\\
            % \(\uu_i^{m} \gets \uu_i^{m-1} + \rho \left( \w_i^{m} - \tilde{\h}_i^{m} \right)\)\;
            % \(\cdots\)\\
            % % \,\\
        }
    }
\end{algorithm}

@@

\section{Evaluation}
\label{sec:simulations}
\subsection[]{Communication cost}
\label{sec:transcost}
\begin{table}[t]
    \centering
    \begin{tabular}{ |l|l|l| } 
        \hline
        & Transmit & Receive \\
        \hline\hline
        (1) Full & \(\mathcal{O}(M-1)\) & \(\mathcal{O}(M-1)\) \\
        \hline
        (2) Broadcast & \(\mathcal{O}(1)\) & \(\mathcal{O}(M-1)\) \\ 
        \hline
        (3) Neigborhood & \(\mathcal{O}(\bar{N}_i K)\) & \(\mathcal{O}(\bar{N}_i K)\) \\ 
        \hline
    \end{tabular}
    \caption[]{Comparison of communication cost.}
    \label{tab:transcost:table}
\end{table}
To describe the communication cost within the network, we adopt the Big-\(\mathcal{O}\) notation, similar to the complexity analysis of algorithms.
We consider the communication of a variable from one node to another as \(\mathcal{O}(1)+\mathcal{O}(1)\), where the two terms represent transmit and receive operations, respectively, as both require computational resources at a node.
The most useful measure in the case of this work is to analyze the cost per node.
For this we define an average neighborhood size \(\bar{N}_i = \frac{1}{M} \sum_{i \in \Mset} N_i\) where \(N_i = | \Nset_i |\) is the cardinality/size of the neigborhood set for node \(i\).
We compare the following communication schemes:
\begin{itemize}
    \itemsep-0.2em
    \item[(1)] A fully connected network, i.e., all nodes are linked with all other nodes by direct links, \(\Nset_i = \Mset\).
    \item[(2)] The nodes employ broadcasting, i.e., the nodes transmit to all other nodes without direct links.
    \item[(3)] The nodes only communicate with neighboring nodes, \(\Nset_i \subset \Mset\).
\end{itemize}
For (1), the cost increases quadratically with increasing network size \(M\), both in transmit and receive operations, whereas for (2), the quadratic increase happens only at the latter.
Using communication scheme (3), as in \cite{yuDistributedBlindSystem2014,liuDistributedBlindIdentification2016} and proposed in this paper, cost in proportional to neighborhood size and iteration count \(K\).
Refer to \autoref{tab:transcost:table} for the collected orders of communication cost.

@@

\subsection[]{Simulation}
\begin{figure}[t]
    \centering
    \input{simulations/plots/icassp2023-static-time/icassp2023-static-time.pgf}
    \caption[]{Median NPM of 30 randomized runs over time for optimal case where all node information is available (fully connected network/broadcasting), fixed values of \(\gamma_i\) with \(K \in \{1,10\}\), adaptive \(\gamma_i\) with \(K=1\). The adaptive method yields results very close to the optimal method. A moving average filter was applied to result curves for better readability (compare transparent and opaque lines).}
    \label{fig:simulations:NPMtime}
\end{figure}
\begin{figure}[t]
    \centering
    \input{simulations/plots/icassp2023-static/icassp2023-static.pgf}
    \caption[]{Mean over time of median NPM of 30 randomized runs after convergence for varied values for fixed \(\gamma_i\) on the interval \([0.0, 0.04]\) and \(K \in \{1,2,10\}\). Results for optimal method, and adaptive \(\gamma_i\) included for comparison.}
    \label{fig:simulations:avgNPMgamma}
\end{figure}
To evaluate the effectiveness of the proposed method, simulations were run.
For this we define the error measure, the normalized projection misalignment
\begin{equation}
    \begin{aligned}
        \text{NPM}(m) &= 20\,\log_{10} \frac{\left\| \mtxb{e} \right\|}{\left\|\h(m)\right\|},\\
        \text{with}\quad \mtxb{e} &= \h(m) - \frac{\h^\T (m) \h_{\text{t}}}{\h_{\text{t}}^\T \h_{\text{t}}}\h(m),
    \end{aligned}
\end{equation}
commonly used in BSI to compare estimated \(\h\) and true \(\h_t\).
The simulation setup is a network with \(M=5\) nodes arranged in a ring topology, each node having \(|N_i|=2\) neighbors.
The input signal is zero-mean white Gaussian noise (WGN), the impulse responses of length \(L=16\) are zero-mean WGN as well, and at each channel, independent WGN is added with \(\text{SNR}=10\,\text{dB}\).
The norms of the impulse responses are scaled to random values from the uniform distribution \(\mathcal{U}_{[0.5,2.0]}\).
A comparison is made between the following cases:
\begin{itemize}
    \itemsep-0.2em
    \item[(a)] all information is available for global norm computation ("optimal" in \autoref{fig:simulations:NPMtime} \& \autoref{fig:simulations:avgNPMgamma}),
    \item[(b)] neighborhood communication with fixed \(\gamma_i \in [0.0, 0.4]\) and \(K \in \{1,2,10\}\),
    \item[(c)] neighborhood communication with adaptive \(\gamma_i\) \eqref{eq:adaptivenormest:adaptivegamma} and \(K=1\).
\end{itemize}
\autoref{fig:simulations:NPMtime} shows the median of 30 Monte-Carlo runs.
We can observe that when fixing \(\gamma_i\), a tradeoff between convergence speed and steady-state error arises at low iteration counts (cf. also \autoref{fig:simulations:avgNPMgamma}).
With \(K=10\), the performance comes close to the optimal case at the cost of additional communication between nodes.
Compare that to the case with adaptive \(\gamma_i\) where even at \(K=1\), i.e., no additional communication for the recursive estimation scheme, convergence speed and steady-state error are close to the optimal case.

\begin{todo}
    TODO: dynamic scenario: add comparisons with optimal and static \(\gamma\) case, and describe in text, add figure captions, extend conclusion
\end{todo}

\begin{figure}[t]
    \centering
    \input{simulations/plots/icassp2023-dynamic-time/icassp2023-dynamic-time.pgf}
    \caption[]{}
    \label{fig:simulations:NPMtimedyn}
\end{figure}
\begin{figure}[t]
    \centering
    \input{simulations/plots/icassp2023-dynamic-norm/icassp2023-dynamic-norm.pgf}
    \caption[]{}
    \label{fig:simulations:normtime}
\end{figure}
\begin{figure}[t]
    \centering
    \input{simulations/plots/icassp2023-dynamic-norms/icassp2023-dynamic-norms.pgf}
    \caption[]{}
    \label{fig:simulations:normstime}
\end{figure}
